---
title: "INN Dashboard Sample"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---


```{r global, include=FALSE}
library(shiny)
library(tidyverse)
library(sf)
library(leaflet)
library(pals)
library(RColorBrewer)
library(DT)
library(highcharter)
library(tigris)
options(tigris_use_cache = T)
library(rmapshaper)

# Read in data and subset for the selected area
load("data/data.rda")

# Define the list of available metros
metrolookup <- sort(unique(na.omit(metroshape$NAME)))
lookup <- sort(unique(na.omit(full_mapdat$statecode)))

# Define colors for mapping
set.seed(12345)
alphabet_pal <- alphabet()
newPal <- grDevices::colorRampPalette(colors = alphabet_pal)(25)
names(newPal) <- levels(factor(full_mapdat$myclus))
factpal <- colorFactor(newPal, full_mapdat$myclus)

# Define map popup labels
# Define map popup labels
full_mapdat$labels <- paste0(
  "<strong> ZIP: </strong> ",
  full_mapdat$zcta, "<br/>",
  "<strong> Cluster: </strong> ",
  full_mapdat$myclus, "<br/>",
  "<strong> Population: </strong> ",
  full_mapdat$pop_2021, "<br/>",
  "<strong> Income: </strong> $ ",
  full_mapdat$medHHincome_2021
) %>% 
  lapply(HTML)


```

Sidebar {.sidebar}
=================================================

### National Market Segments

The segments are based on various set of demographic, socioeconomic status variables at the zip code level. Please select a state from below. You can freely browse the map to explore specific Zip Codes.

```{r}
# Define metro area
# selectInput("state", label = "Select a State", choices = lookup,
#             selected = "IL")
selectInput("metroname", label = "Select a Metro Area", choices = metrolookup,
            selected = "Chicago-Naperville-Evanston, IL")
```

Explore Market Segments
=================================================

Column
-------------------------------------------------

### Map

```{r}
# Reactive subset
# sub_mapdat <- reactive({
#   full_mapdat %>% 
#     filter(statecode == input$state)
# })

sub_mapdat <- reactive({
  metroshape %>% 
    filter(NAME == input$metroname)
})

# Create map
output$map <- renderLeaflet({
  bbox <- st_bbox(sub_mapdat()) %>% as.vector()
  # Create leaflet object
  leaflet() %>% 
    addProviderTiles("CartoDB.Positron") %>% 
    addPolygons(data = full_mapdat,
                fillColor = ~factpal(myclus),
                color = "white",
                weight = 0.5,
                smoothFactor = 0,
                fillOpacity = 0.7,
                layerId = ~zcta,
                label = ~labels,
                group = "fullclus",
                highlight = highlightOptions(
                  color = "cyan",
                  bringToFront = T)) %>% 
    setView(lng = (bbox[1] + bbox[3]) / 2,  # Center of the bounding box
            lat = (bbox[2] + bbox[4]) / 2,  # Center of the bounding box
            zoom = 10) %>% 
    addLegend(pal = factpal, values = full_mapdat$myclus, title = "Segments", group = "fullclus")
})


# Click event for the map (used to generate table/chart)
click_zip <- eventReactive(input$map_shape_click, {

  x <- input$map_shape_click

  y <- x$id

  return(y)
})

# Observe clicks
observeEvent(input$map_shape_click, {
  # Create reactive storing layerid of clicked zips
  zip_ids <- reactive({
    eventdata <- input$map_shape_click
    if (is.null(eventdata)) {
      return(NULL) # do nothing
    } else {
      zips <- eventdata$id
      return(zips)
    }
  })
  # Create leaflet proxy
  
  # Subset the clicked zipcode
  sub <- filter(full_mapdat, zcta %in% zip_ids())
  # Get boundaries for selected zip
  bbox2 <- st_bbox(sub)
  xrange <- bbox2$xmax - bbox2$xmin
  yrange <- bbox2$ymax - bbox2$ymin
  bbox2[1] <- bbox2[1] - xrange
  bbox2[2] <- bbox2[2] - yrange
  bbox2[3] <- bbox2[3] + xrange
  bbox2[4] <- bbox2[4] + yrange
  bbox2 <- bbox2 %>% as.vector()
  
  proxy <- leafletProxy("map")
  # Clear old selection on map, and add new selection
  proxy %>%
    clearGroup(group = "sub") %>%
    addPolygons(data = sub, fill = F, color = "#FFFF00",
                opacity = 1, group = "sub", weight = 5) %>% 
    fitBounds(lng1 = bbox2[1],
              lat1 = bbox2[2],
              lng2 = bbox2[3],
              lat2 = bbox2[4])

})

zip_data <- reactive({
  # Fetch data for the clicked zip
  return(filter(full_mapdat, zcta == click_zip()))
})

leafletOutput("map")



```

### Digital Profile Table

```{r}
# Create reactive subset data
metro_sub <- reactive({
  submsa <- zip_data()$msa
  metsub <- full_mapdat %>% 
    filter(msa == submsa) %>% 
    select(zcta, myclus, medHHincome_2021, popPctCh, bbsub_none, bbsub_onlycell, dv_onlycell) %>% 
    st_drop_geometry() %>%
    mutate(across(c(starts_with("bbsub"), popPctCh, dv_onlycell), ~ .x / 100))
  return(metsub)
})

# output$dt <- renderDataTable({
#   
#   dt <- full_mapdat %>% 
#     select(zcta, myclus, medHHincome_2021, popPctCh, bbsub_none, bbsub_onlycell, dv_onlycell) %>% 
#     st_drop_geometry() %>% 
#     mutate(across(c(starts_with("bbsub"), popPctCh, dv_onlycell), ~ .x / 100))
#   
#   datatable(dt, rownames = F,
#             colnames = c("Zip Code", "Segment", "Income", "Population Change", "No Broadband Subscription",
#                                                  "Only Mobile Broadband Subscriptions", "Only Smartphone"),
#             options = list(pageLength = 10)) %>% 
#     formatPercentage(4:7) %>% formatCurrency(3, digits = 0)
#   
# })

output$dt <- renderDataTable({
  
  # dt <- metro_sub() %>% 
  #   select(zcta, myclus, medHHincome_2021, popPctCh, bbsub_none, bbsub_onlycell, dv_onlycell) %>% 
  #   st_drop_geometry() %>%
  #   mutate(across(c(starts_with("bbsub"), popPctCh, dv_onlycell), ~ .x / 100))
  # 
  datatable(metro_sub(), rownames = F,
            colnames = c("Zip Code", "Segment", "Income", "Population Change", "No Broadband Subscription",
                                                 "Only Mobile Broadband Subscriptions", "Only Smartphone"),
            options = list(pageLength = 10)) %>% 
    formatPercentage(4:7) %>% formatCurrency(3, digits = 0)
  
})

dataTableOutput("dt")

```

Column
-------------------------------------------------

### Race/Ethnic Profile, selected Zip code, Segment, and Metro Average (click on the map to show chart)

```{r}
output$raceplot <- renderHighchart({
  
  zd <- zip_data()
  
  # Set up metro data
  metro_data <- full_mapdat %>% st_drop_geometry() %>% 
    filter(metro_group == zd$metro_group) %>% 
    group_by(metro_group) %>% 
    summarise(across(c(starts_with("race") & ends_with("2021"), lessthan9_2021, highschool_2021,
                       college_2021, bachelor_2021, graduate_2021), ~ round(mean(.x, na.rm = T), 1)))
    
  # Set up segment data
  clus_data <- full_mapdat %>% st_drop_geometry() %>% 
    filter(myclus == zd$myclus) %>% 
    group_by(myclus) %>% 
    summarise(across(c(starts_with("race") & ends_with("2021"), lessthan9_2021, highschool_2021,
                       college_2021, bachelor_2021, graduate_2021), ~ round(mean(.x, na.rm = T), 1)))
    
  
  chart <- highchart() %>% 
    hc_chart(type = "column") %>% 
    hc_xAxis(categories = c("White", "Black", "Hispanic", "Asian", "Native", "Pacific-Islander", "Other"),
             title = list(text = 'Race/ethnicity')) %>% 
    hc_yAxis(title = list(text = "%")) %>% 
    hc_plotOptions(series = list(dataLabels = list(enabled = T))) %>% 
    hc_add_series(name = paste0("Zip Code = ", as.character(zd$zcta)),
                  data = c(zd$race_white_2021,
                           zd$race_black_2021,
                           zd$race_hisp_2021,
                           zd$race_asian_2021,
                           zd$race_native_2021,
                           zd$race_pacific_2021,
                           zd$race_other_2021)) %>% 
    hc_add_series(name = paste0("Metro Group Average ", as.character(zd$metro_group)),
                  data = c(metro_data$race_white_2021,
                           metro_data$race_black_2021,
                           metro_data$race_hisp_2021,
                           metro_data$race_asian_2021,
                           metro_data$race_native_2021,
                           metro_data$race_pacific_2021,
                           metro_data$race_other_2021)) %>% 
    hc_add_series(name = paste0("Segment = ", as.character(zd$myclus)),
                  data = c(clus_data$race_white_2021,
                           clus_data$race_black_2021,
                           clus_data$race_hisp_2021,
                           clus_data$race_asian_2021,
                           clus_data$race_native_2021,
                           clus_data$race_pacific_2021,
                           clus_data$race_other_2021)) %>% 
    hc_title(text = "Racial/Ethnic Profile Comparison (%)", align = "left") %>% 
    hc_add_theme(hc_theme_smpl()) %>% 
    hc_tooltip(enabled = F)
  
  chart
  
})

highchartOutput("raceplot")

```

### Educational Attainment Profile, selected Zip code, Segment, and Metro Average (click on the map to show chart)

```{r}
output$eduplot <- renderHighchart({
  
  zd <- zip_data()
  
  # Set up metro data
  metro_data <- full_mapdat %>% st_drop_geometry() %>% 
    filter(metro_group == zd$metro_group) %>% 
    group_by(metro_group) %>% 
    summarise(across(c(starts_with("race") & ends_with("2021"), lessthan9_2021, highschool_2021,
                       college_2021, bachelor_2021, graduate_2021), ~ round(mean(.x, na.rm = T), 1)))
    
  # Set up segment data
  clus_data <- full_mapdat %>% st_drop_geometry() %>% 
    filter(myclus == zd$myclus) %>% 
    group_by(myclus) %>% 
    summarise(across(c(starts_with("race") & ends_with("2021"), lessthan9_2021, highschool_2021,
                       college_2021, bachelor_2021, graduate_2021), ~ round(mean(.x, na.rm = T), 1)))
    
  
  chart <- highchart() %>% 
    hc_chart(type = "column") %>% 
    hc_xAxis(categories = c("Less than High", "Highschool", "Some College", "Bacehlor", "Post Graduate"),
             title = list(text = 'Educational Attainment')) %>% 
    hc_yAxis(title = list(text = "%")) %>% 
    hc_plotOptions(series = list(dataLabels = list(enabled = T))) %>% 
    hc_add_series(name = paste0("Zip Code = ", as.character(zd$zcta)),
                  data = c(zd$lessthan9_2021,
                           zd$highschool_2021,
                           zd$college_2021,
                           zd$bachelor_2021,
                           zd$graduate_2021)) %>% 
    hc_add_series(name = paste0("Metro Group Average ", as.character(zd$metro_group)),
                  data = c(metro_data$lessthan9_2021,
                           metro_data$highschool_2021,
                           metro_data$college_2021,
                           metro_data$bachelor_2021,
                           metro_data$graduate_2021)) %>% 
    hc_add_series(name = paste0("Segment = ", as.character(zd$myclus)),
                  data = c(clus_data$lessthan9_2021,
                           clus_data$highschool_2021,
                           clus_data$college_2021,
                           clus_data$bachelor_2021,
                           clus_data$graduate_2021)) %>% 
    hc_title(text = "Educational Attainment Level Comparison (%)", align = "left") %>% 
    hc_add_theme(hc_theme_smpl()) %>% 
    hc_tooltip(enabled = F)
  
  chart
  
})

highchartOutput("eduplot")

```

About Segmentation Scheme
=================================================

**Expalantion about the segmentation scheme**
